Argennon: A Scalable Smart Contract Platform

### 2021-07-03 tex/sections/blocks.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

Having the previous AVM state, the transaction list and the clustering directives of a block, a node can calculate
commitments to the staking, method and heap databases of the current block by emulating the AVM execution. If the
node can obtain the previous block information from a trusted source, it does not need to have a trusted local
copy of the AVM state,
and it can reliably retrieve the required storage pages from a ZK-EDB server. We call this type of block verification
\emph{conditional} block validation. This validation is conditional because the validity of the current block is
conditioned on the validity of the previous block.

@@

Argennon block certificates constitute an ordered sequence based on the order of blocks they certify. If we show
the certificate of committee $C$ for the $i$-th block\footnote{note that the $i$-th block certificate is not
necessarily the certificate of the $i$-th block.} with $\C_i$, and the set of signers
with $S_i$, then the block certificate $\C_i$ can be considered as a tuple:
\begin{equation}
    \C_i=(\sigma_i,\ C-S_{i})\label{eq:cert}\ ,
\end{equation}
where $\sigma_i$ is the aggregate signature issued by $S_i$.

@@

Alternately we can use $\APK_{i-1}$ to calculate the aggregate public key:
\begin{equation}
    \APK_i=\APK_{i-1}\APK_{S_i-S_{i-1}}\APK_{S_{i-1}-S_i}^{-1}\ .\label{eq:aggPK-2}
\end{equation}

>>>>

Having the previous AVM state, the transaction list and the clustering directives of a block, a node can calculate
commitments to the staking, method and heap databases of the current block by emulating the AVM execution. If the
node can obtain the previous block commitments from a trusted source, it does not need to have a trusted local
copy of the AVM state,
and it can reliably retrieve the required storage pages from a ZK-EDB server. We call this type of block verification
\emph{conditional} block validation, since the validity of the current block is
conditioned on the validity of the previous block.

@@

Argennon block certificates constitute an ordered sequence based on the order of blocks they certify. If we show
the $i$-th certificate\footnote{note that the $i$-th certificate is not
necessarily the certificate of the $i$-th block.} of committee $C$ with $\C_i$, and the set of signers
with $S_i$, then the block certificate $\C_i$ can be considered as a tuple:
\begin{equation}
    \C_i=(\sigma_i,\ C-S_{i})\label{eq:cert}\ ,
\end{equation}
where $\sigma_i$ is the aggregate signature issued by $S_i$.

@@

Alternately, we can use $\APK_{i-1}$ to calculate the aggregate public key:
\begin{equation}
    \APK_i=\APK_{i-1}\APK_{S_i-S_{i-1}}\APK_{S_{i-1}-S_i}^{-1}\ .\label{eq:aggPK-2}
\end{equation}

### 2021-07-03 tex/sections/consensus.tex
Add incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The credibility of a block of the Argennon blockchain is determined based on the certificates it has received
from different sets of users, knows as committees. There are two primary types of certificate committees in
Argennon: the committee of \emph{delegates} and the committee
of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.

@@

Every \texttt{committeeLifeTime} number of blocks, randomly $m$ committees are selected from
validators, in a way that the total stake of members of each committee is approximately equal, and every
account is a member of \textbf{at least} one committee.

@@

A block certificate issued by some members of a validators' committee is considered valid, if according to
the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
the stake values based on the previous block a malicious committee can select the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of certificate signers is higher than $3/4$ of the total stake
    of the committee members that have \texttt{online} status.
\end{itemize}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not received
a certificate from the validators yet.

@@

So in Argennon, the block validation by committees is performed in parallel, and validators
do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
certificate if the certificate of the previous block has not been published yet. This ensures that an invalid
fork made by malicious delegates will not receive any certificates from validators.

@@

If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
be very low, even zero. Therefore, the status block should not be certified only by online stake.} After applying
the transactions of the status block, the total online stake of validators
must go higher than \texttt{minOnlineStake} threshold.

@@

The recovery protocol is a resilient protocol designed for recovering the Argennon blockchain from critical situations.
In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
during the lifetime of the Argennon blockchain.

@@

The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
from the validators.
This fork needs to be confirmed by the validators and can not discard any blocks that has been already certified
by validators.
A valid fork request is an unexpired request signed by more than half of the
total \texttt{online} stake of the validators.

@@

As long as more than half of the total online stake of every committee of validators is controlled
by honest users, a malicious committee of delegates can not use the emergency forking procedure to discard blocks
that have a validators' certificate.

@@

The agreement protocol consists of two phases: the \emph{voting phase}, which selects a single proposal
and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current session ends and a new
agreement session starts. All votes and messages are tagged in a way that the messages of one session can
not be used in another session.

@@

When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
confirmation phase could be skipped. In that case, the confirmation of the fork block by the committee of validators
acts like the confirmation phase.

@@

To do so, first the validator activates the censorship resilient mode of his networking module, then he checks the
validity of the blocks in his blockchain that do not have a validators' certificate and determines the
last valid block of his blockchain.

@@

In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
of stake the user has in the system. Conventionally in these systems, a user's stake is considered to be equal with the
amount of native system tokens, the user has "staked" in the system. A user stakes his tokens by locking them in
his account or a staking account for some period of time, and during this time he will not be able to transfer
his tokens.

Unfortunately, there is a subtle problem with this approach. It is not clear that in a real world economic system
how much of the main currency of the system can be locked and kept out of the circulation indefinitely. It seems that
this amount for currencies like US dollar, is quite low comparing to the total market cap of the currency.
This means that for a real world currency this type of staking mechanisms will result in putting the
fate of the system in the hands of the owners of a small fraction of the total supply of a currency.

To mitigate this problem, Argennon uses a hybrid approach for estimating the stake of a user.
Every \texttt{stakingDuration} blocks, which is called a \emph{staking period}, Argennon calculates
a \emph{trust value} for each user. The user's stake
at time step \(t\), is estimated based on the user's trust value and his ARG balance:
\begin{equation}
    S_{u,t} = \min (B_{u,t}, Trust_{u,k})\ ,\label{eq:stake}
\end{equation}
where:
\begin{itemize}
    \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
    \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
    \item \(Trust_{u,k}\) is an estimated trust value for user \(u\) at staking period \(k\).
\end{itemize}

>>>>

The credibility of a block of the Argennon blockchain is determined by the certificates it has received
from different sets of users, knows as committees. There are two primary type of certificate committees in
Argennon: the committee of \emph{delegates} and the committee
of \emph{validators}. Argennon has \emph{one} committee of delegates and $m$ committees of validators.

@@

Every \texttt{committeeLifeTime} number of blocks, randomly $m$ committees are selected from
validators, in a way that the total stakes of committees are almost equal, and every
account is a member of \textbf{at least} one committee.

@@

A block certificate issued by some members of a validators' committee is considered valid, if according to
the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
the stake values based on the previous block a malicious committee can select the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} fraction
    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than $3/4$ of the total stake
    of the committee members that have \texttt{online} status.
\end{itemize}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not
yet received a certificate from the validators.

@@

So in Argennon, the block validation by committees is performed in parallel, and validators
do not wait for seeing the certificate of the previous block validators to start transaction validation. On the
other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
certificate if the certificate of the previous block has not been published yet. This ensures that an invalid
fork made by malicious delegates will not receive any certificates from validators.

@@

If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
be very low. Therefore, the status block should not be certified only by online stake.} After applying
the transactions of the status block, the total online stake of all committees of validators
must go higher than \texttt{minOnlineStake} threshold.

@@

The recovery protocol is a resilient protocol designed for recovering the Argennon blockchain from critical situations.
In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability,
and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
during the lifetime of the Argennon blockchain.

@@

The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
from the validators.
This fork needs to be confirmed by validators and can not discard any blocks that has been already certified
by validators.
A valid fork request is an unexpired request signed by more than half of the
total \texttt{online} stake of the validators.

@@

As long as more than half of the total online stake of every committee of validators is controlled
by honest users, a malicious committee of delegates can not use the emergency forking procedure to discard blocks
that have a certificate from validators.

@@

The agreement protocol consists of two phases: the \emph{voting phase}, which selects a single proposal
and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current agreement session ends and a
new session starts. All votes and messages are tagged in a way that the messages of one session can
not be used in another session.

@@

When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
confirmation phase could be skipped. In that case, the confirmation of the fork block by the appropriate committee
of validators acts like the confirmation phase.

@@

To do so, first the validator activates the censorship resilient mode of his networking module, then he checks the
validity of the blocks that do not have a validators' certificate and determines the
last valid block of his blockchain.

@@

In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
of stake the user has in the system. Conventionally in these systems, a user's stake is considered to be equal with the
amount of native system tokens, he has "staked" in the system. A user stakes his tokens by locking them in
his account or a separate staking account for some period of time. During this time, he will not be able to transfer
his tokens.

Unfortunately, there is a subtle problem with this approach. It is not clear that in a real world economic system
how much of the main currency of the system can be locked and kept out of the circulation indefinitely. It seems that
this amount for currencies like US dollar, is quite low comparing to the total market cap of the currency.
This means that for a real world currency this type of staking mechanisms will result in putting the
fate of the system in the hands of the owners of a small fraction of the total supply.

To mitigate this problem, Argennon uses a hybrid approach for estimating the stake of a user.
Every \texttt{stakingDuration} blocks, which is called a \emph{staking period}, Argennon calculates
a \emph{trust value} for each user.

The user's stake
at time step \(t\), is estimated based on the user's trust value and his ARG balance:
\begin{equation}
    S_{u,t} = \min (B_{u,t}, Trust_{u,k})\ ,\label{eq:stake}
\end{equation}
where:
\begin{itemize}
    \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
    \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
    \item \(Trust_{u,k}\) is an estimated trust value for user \(u\) at staking period \(k\).
\end{itemize}

### 2021-07-03 tex/sections/elements.tex
Add incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

\note{Explicit key registration enables Argennon to decouple cryptography from blockchain design. In this way,
    if the used cryptographic algorithms turn out to be insecure for some reason, for example because
    of the introduction of quantum computers, the cryptographic algorithms could be easily upgraded.}

@@

\begin{itemize}
    \item \texttt{avmCall}: invokes a method from an
    AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
    assets, including ARGs, is done by these transactions.
    \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
    on.
    \item \texttt{unInstallApp}: removes an AVM smart contract.
    \item \texttt{updateApp}: updates an AVM smart contract (if allowed).
\end{itemize}

@@

Also, Argennon transactions are required to specify what heap or code area addresses they will access. This will
enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. A transaction
that tries to access a memory location that is not in its access lists, will be rejected.
Users could use off-chain \emph{smart contract oracles} to predict the list of memory locations their transactions need.

A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM memory and can emulate AVM
execution without accessing a ZK-EDB server. Smart contract oracles can be used for reporting useful information about
Argennon transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
and so on.

@@

Every Argennon transaction is required to pay two types of fees: execution fee and storage fee. A transaction pays
its fees by providing digital signatures of one or more accounts authorizing the transfer of the amount of fee in
ARG from one or more accounts to the fee sink accounts.
\note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
were not used in the final emulation. This will incentivize users to report the resource usage of their
transactions more accurately.}

>>>>

\note{Explicit key registration enables Argennon to decouple cryptography from the blockchain design. In this way,
    if the cryptographic algorithms used become insecure for some reason, for example because
    of the introduction of quantum computers, they could be easily upgraded.}

@@

\begin{itemize}
    \item \texttt{avmCall}: invokes a method from an
    AVM smart contract. Users interact with AVM smart contracts using this method of the root smart contract.
    Transferring all assets, including ARG, is done by these transactions.
    \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
    on.
    \item \texttt{unInstallApp}: removes an AVM smart contract.
    \item \texttt{updateApp}: updates an AVM smart contract (if allowed).
\end{itemize}

@@

Also, Argennon transactions are required to specify what heap or code area addresses they will access. This will
enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. A transaction
that tries to access a memory location that is not in its access list, will be rejected.
Users could use off-chain \emph{smart contract oracles} to predict the list of memory locations their transactions need.

A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM storage and can emulate AVM
execution without accessing a ZK-EDB server. Smart contract oracles can be used for reporting useful information about
Argennon transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
and so on.

@@

Every Argennon transaction is required to pay two types of fees: execution fee and storage fee. A transaction pays
its fees by providing digital signatures of one or more accounts, authorizing the transfer of the amount of fee in
ARG from one or more accounts to the fee sink accounts.

\note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
were not used in the final emulation. This will incentivize users to report the resource usage of their
transactions more accurately.}
